Multi-target detection with application to cryo-electron microscopy

### 2019-03-08 inverse_problems/inverse_problems.tex
replying to Amit's comments

Similarly to the homogeneous well-separated model, the observed autocorrelations under the Poisson model identify $x,\gamma$ and $\sigma^2$ uniquely.
Opposed to Proposition~\ref{prop:gamma_sigma}, these quantities can be computed explicitly. 
In Section~\ref{sec:proof_uniqueness_poisson} we prove the following result:
%
%We first note that $\gamma a_x^1$, $\gamma a_x^2[\ell] + \sigma^2 \ell[\ell]$, and $\gamma a_x^3[\ell_1,\ell_2]$ can be computed directly form the observed autocorrelations~\eqref{eq:mean_micrograph2},~\eqref{eq:ac2_micrograph2} and~\eqref{eq:ac3_micrograph2}.
% Indeed, recovering  $\gamma a_x^1$ and $\gamma a_x^2[\ell] + \sigma^2 \ell[\ell]$ is immediate from \eqref{eq:mean_micrograph2} and \eqref{eq:ac2_micrograph2}, and $\gamma a_x^3[\ell_1,\ell_2]$ then follows from
%%
%\begin{align}
%\gamma a_x^3[\ell_1,\ell_2]
%= a_y^3[\ell_1,\ell_2] - (a_y^1)^3
%- a_y^1 \cdot \big(a_y^2[\ell_1]  
%+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(a_y^1)^2 \big).
%\end{align}
%Using these expressions, we prove the following in Section~\ref{sec:proof_uniqueness_poisson}.
\begin{proposition} \label{prop:uniqueness_poisson}
Under the homogeneous Poisson model, the signal $x$, the noise level $\sigma^2$ and the Poisson parameter $\gamma$ are identified uniquely. 
In particular, \TODO{Shouldn't it be $L^2$ for $\gamma$?}
	\begin{align}
\gamma &= L \frac{(\gamma a_y^1) (a_y^2[1] - (a_y^1)^2)}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=2}^{L-1} \gamma a_x^3[\ell,\ell+1]}, \\
\sigma^2 &= a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2,
\end{align}
where 
\begin{align}
\gamma a_x^3[\ell_1,\ell_2]
&= a_y^3[\ell_1,\ell_2] - (a_y^1)^3
- a_y^1 \cdot \big(a_y^2[\ell_1]  
+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(a_y^1)^2 \big).
\end{align}

@@

Hence, if $\sigma$ is known we collect
\begin{align*}
	1 + L + \frac{(L+1)(L+2)}{2}-2 = \frac{1}{2} L (L+5)
\end{align*}
equations, while if it unknown and we choose not to estimate it, then we collect
\begin{align*}
	1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
equations in total. Of course, there may be redundancy in these equations: we aim only to provide an upper bound. %under the assumption that these equations are independent.

@@

In this section, we present three numerical experiments. In the first two experiments, we consider the heterogeneous model~\eqref{eq:model}. First, with a fixed noise level, we show how the estimation quality improves as the length of the observation grows. Second, we explore how many signals can be recovered as a function of their length, in an infinite data regime where effects of the noise have been averaged out. In the last experiment, we extend our model to 2-D signals.
The code for all experiments is available at \url{https://github.com/PrincetonUniversity/BreakingDetectionLimit}. \TODO{Need to upload the code}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=1\linewidth]{heterogeneous_progressive_n12300000000_466300}
	\caption{Experiment described in Section~\ref{sec:XP1}. For a fixed noise level $\sigma = 3$ and a fixed set of $K = 3$ signals of length $L = 21$, an observation $y$ of length $N = 12.3 \cdot 10^9$ is generated according to~\eqref{eq:model} in the well-separated model, with fixed d occurrence probabilities. Each rows corresponds to one of the signals. The last column shows evolution of the relative root mean squared error in estimating each signal, as a longer and longer subset of $y$ is observed. Red dots mark the three snapshots that are illustrated in columns 1--3: red signals are the ground truth and blue signals are the estimators.}
	\label{fig:1Dheterosignals}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.7\linewidth]{KLXP/XP1}
	\caption{In the $N \to \infty$ regime (access to exact autocorrelations, excluding biased entries) and with known uniform densities, it seems that $K$ up to $\sqrt{L}$ (red curve) i.i.d.\ Gaussian signals of length $L$ can be recovered from the known moments. CPU time is in seconds. Strictly above red dots, recovery is impossible because the number of unknowns exceeds the number of computed autocorrelation; see Section~\ref{sec:heterogeneity}. Similarly to~\cite[Fig.~4.1]{boumal2017heterogeneous}, this experiment suggests a possible statistical-computational gap.}
	\label{fig:KLXP}
\end{figure}

@@

\section*{Acknowledgment} \TODO{To check}
The authors  thank Ayelet Heimowitz, Joe Kileel, Ti-Yen Lan and Amit Moscovich  for helpful discussions.
The research was partially supported by Award Number R01GM090200 from the NIGMS, FA9550-17-1-0291 from AFOSR, Simons Foundation Math+X Investigator Award, and the Moore Foundation Data-Driven Discovery Investigator Award.
NB is partially supported by NSF award DMS-1719558.

>>>>

Similarly to the homogeneous well-separated model, the observed autocorrelations under the Poisson model identify $x,\gamma$ and $\sigma^2$ uniquely.
Opposed to Proposition~\ref{prop:gamma_sigma}, these quantities can be computed explicitly. 
In Section~\ref{sec:proof_uniqueness_poisson} we prove the following result:

\begin{proposition} \label{prop:uniqueness_poisson}
Under the homogeneous Poisson model, the signal $x$, the noise level $\sigma^2$ and the Poisson parameter $\gamma$ are identified uniquely. 
In particular,
	\begin{align}
\gamma &= L \frac{(\gamma a_y^1) (a_y^2[1] - (a_y^1)^2)}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=2}^{L-1} \gamma a_x^3[\ell,\ell+1]}, \\
\sigma^2 &= a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2,
\end{align}
where 
\begin{align}
\gamma a_x^3[\ell_1,\ell_2]
&= a_y^3[\ell_1,\ell_2] - (a_y^1)^3
- a_y^1 \cdot \big(a_y^2[\ell_1]  
+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(a_y^1)^2 \big).
\end{align}

@@

Hence, if $\sigma$ is known we collect
\begin{align*}
	1 + L + \frac{(L+1)(L+2)}{2}-2 = \frac{1}{2} L (L+5)
\end{align*}
equations, while if it is unknown and we choose not to estimate it, then we collect
\begin{align*}
	1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
equations in total. Of course, there may be redundancy in these equations: we aim only to provide an upper bound. %under the assumption that these equations are independent.

@@

In this section, we present three numerical experiments. In the first two experiments, we consider the heterogeneous model~\eqref{eq:model}. First, with a fixed noise level, we show how the estimation quality improves as the length of the observation grows. Second, we explore how many signals can be recovered as a function of their length, in an infinite data regime where effects of the noise have been averaged out. In the last experiment, we extend our model to 2-D signals.
The code for all experiments is available at \url{https://github.com/PrincetonUniversity/BreakingDetectionLimit}.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=1\linewidth]{heterogeneous_progressive_n12300000000_466300}
	\caption{Experiment described in Section~\ref{sec:XP1}. For a fixed noise level $\sigma = 3$ and a fixed set of $K = 3$ signals of length $L = 21$, an observation $y$ of length $N = 12.3 \cdot 10^9$ is generated according to~\eqref{eq:model} in the well-separated model, with fixed  occurrence probabilities. Each row corresponds to one of the signals. The last column shows evolution of the relative root mean squared error in estimating each signal, as a longer and longer subset of $y$ is observed. Red dots mark the three snapshots that are illustrated in columns 1--3: red signals are the ground truth and blue signals are the estimators.}
	\label{fig:1Dheterosignals}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.7\linewidth]{KLXP/XP1}
	\caption{In the $N \to \infty$ regime (access to exact autocorrelations, excluding biased entries) and with known uniform densities, it seems that $K$ up to $\sqrt{L}$ (red curve) i.i.d.\ Gaussian signals of length $L$ can be recovered from the known moments. CPU time is in seconds. Strictly above red dots, recovery is impossible because the number of unknowns exceeds the number of computed autocorrelations; see Section~\ref{sec:heterogeneity}. Similarly to~\cite[Fig.~4.1]{boumal2017heterogeneous}, this experiment suggests a possible statistical-computational gap.}
	\label{fig:KLXP}
\end{figure}

@@

\section*{Acknowledgment} 
The authors  thank Ayelet Heimowitz, Joe Kileel, Ti-Yen Lan and Amit Moscovich  for helpful discussions.
The research was partially supported by Award Number R01GM090200 from the NIGMS, FA9550-17-1-0291 from AFOSR, Simons Foundation Math+X Investigator Award, the Moore Foundation Data-Driven Discovery Investigator Award, and NSF BIGDATA Award IIS-1837992.
NB is partially supported by NSF award DMS-1719558.
